### 2015-05-02 MSDgroupwriteup.tex
modified script to get sparsity upates to tex doc

% \documentclass[12pt]{emulateapj}
 \documentclass[12pt,preprint]{aastex}
\usepackage[margin= 1.0in]{geometry}    % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper} % or letter or a5paper or ... etc
\usepackage{float}
\usepackage{amssymb,amsmath}
\usepackage[]{epsfig,graphicx}
\usepackage{color}
\usepackage{verbatim}
\DeclareGraphicsRule{.tif}{png}{.jpg}{`convert Num1 `dirname Num1`/`basename Num1 .tif`.jpg}
\newcommand{\units}[1]{\ensuremath{\, \mathrm{#1}}}
\newcommand{\myemail}{egrand@astro.umd.edu}
\usepackage{enumitem}
\usepackage{natbib,twoopt}
\newcommand{\degree}{\ensuremath{^\circ}}
\citestyle{apj}
\bibliographystyle{aa}
\usepackage[maxfloats=25]{morefloats}
\usepackage[normalem]{ulem}
\usepackage{hyperref}

\begin{document}
\title{Clustering of Last.fm Music Artists}

 \author{Jordan Rosenblum, Justin Law \& Erin Grand}
 \affil{Data Science Institute, Columbia University, New York, NY 10027}

\date{\today}

\begin{abstract}
This is the abstract. 
\end{abstract}

This is the body.

\href{http://www.dtic.upf.edu/~ocelma/MusicRecommendationDataset/}{Music Recommendation Dataset}

Sparsity of $8.08199453451\times 10^{-5}$

\end{document}